One-arm spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-06 paper/results_3d.tex
update
minor edits

\section{Three-dimensional simulations}
In this section we briefly review results from 3D simulations carried 
out using the ZEUS-MP and PLUTO. The main purpose is to check 
the above results with different numerical codes, and validate  
the 2D approximation.

The 3D disc has radial size
$[r_\mathrm{min},r_\mathrm{max}]=[0.4,10]R_0$ and vertical extent 
$n_H=2$ scale-heights. The resolution is $N_r\times N_\theta\times
N_\phi=256\times32\times256$. Because of the much reduced resolution
compared to 2D, we use a smooth perturbation by setting
$\delta = 10^{-3}$ and $M=1$ in Eq. \ref{randpert}. This corresponds
to a single $m=1$ spiral in the dead zone, which was found to dominate
the 2D simulation.

@@

Snapshots at the end of the simulations are shown in
Fig. \ref{3d_prelim} for both ZEUS-MP and PLUTO. We again find the dead zone eventually
dominated by an $m=1$ spiral. %A Cartesian visualisation of the
% midplane density perturbation associated with the $m=1$ mode is shown in Fig. \ref{pluto_cart}.
The two codes give qualitatively similar results, but a longer PLUTO simulation was
needed to achieve similar mode amplitudes in ZEUS-MP.

%The amplitude of the spiral is smaller than the 2D case above,
%possibily due to lower resolution and/or 3D self-gravity being weaker
%than that in 2D.

\begin{figure}
  \begin{center}
    \subfigure[ZEUS-MP]{
      \includegraphics[scale=0.33]{figures/polarxy_dens015_zeus}
    }
    \subfigure[PLUTO]{
      \includegraphics[scale=0.33]{figures/pdisk_020}
    }
  \end{center}
  \caption{Preliminary 3D simulations using the (a) ZEUS-MP and (b)
    PLUTO. The non-axisymmetric midplane density at the end of the run
    is shown. Here $\psi \equiv \pi/2 - \theta$ is the angular height
    from the midplane.\label{3d_prelim}}   
\end{figure}

% \begin{figure}
%   \includegraphics[width=\linewidth]{figures/pdiskxy_020_m1}
%   \caption{Cartesian visualisation the $m=1$ midplane density in the PLUTO
%     simulation shown in Fig. \ref{3d_prelim}.\label{pluto_cart}}   
% \end{figure}

%\subsection{Angular momentum evolution}

@@

%Note that FARGO does not explicitly conserve angular
%momentum. In addition, there may be an angular momentum flux due to
%self-gravity across disc boundaries since our domain size is finite. 
%Fig. \ref{prelim_angmom} shows the angular momentum evolution in
%the 3D runs.

@@

% \section{One-arm spirals in structured discs}
% Here we examine a 2D FARGO simulation in more detail. The setup is the
% same as that in \S\ref{fargo_fiducial}, except we use a resolution of
% $N_R\times N_\phi = 2048\times4096$ and initialise the disc with a
% single $m=1$ perturbation.

% \subsection{Properties of the $m=1$ dead zone spiral}
% % growth rate, co-rotation radius, pattern speed, propagation diagram

>>>>

\section{Three-dimensional simulations}
In this section we briefly review 3D simulations carried 
out using ZEUS-MP and PLUTO. The main purpose is to check 
the above results with different numerical codes, and validate  
the 2D approximation.

The 3D disc has radial size
$[r_\mathrm{min},r_\mathrm{max}]=[0.4,10]R_0$ and vertical extent 
$n_H=2$ scale-heights. The resolution is $N_r\times N_\theta\times
N_\phi=256\times32\times256$. This corresponds to about $4$ cells per
$H$in radius. 
Because of the much reduced resolution 
compared to 2D, we use a smooth perturbation by setting
$\delta = 10^{-3}$ and $M=1$ in Eq. \ref{randpert}. This corresponds
to a single $m=1$ disturbance in $R\in[R_1,R_2]$, which was found to dominate
the 2D simulation.

@@

Fig. \ref{3d_ampmax} plots the amplitude of the $m=1$ spirals measured
in the ZEUS-MP and PLUTO runs. We also show results from strictly
isothermal discs ($q=0$), which clearly display no growth. %why no high m
                                %modes as in 2D for iso disc?
                                %resolution?
Results from ZEUS-MP appears to be off-set from that obtained using
PLUTO. We suspect this may be due to larger numerical noise in the
ZEUS-MP code. Thus the $m=1$ spiral grows to a larger amplitude by the
end of the run. However, we measure similar growth rates,
\begin{align*}
&\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
&\gamma \simeq 0.0075\Omega_k(R_0) \quad\quad \mathrm{ZEUS-MP}.
\end{align*}
Both are smaller than the 2D simulations above. This may be due to the
low resolutions adopted. 
%growth rate does not depend on pert amp

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_ampmax3d.ps}
  \caption{Evolution of the maximum $m=1$ density in  $r\in[R_1,R_2]$
    measured in the 3D simulations. Both discs with a imposed
    temperature gradient ($q=1$) and a strictly isothermal disc
    ($q=0$) are shown. 
    \label{3d_ampmax}}   
\end{figure}

Visualisations of the 3D simluations are shown in Fig. \ref{3d_prelim} for
both ZEUS-MP and PLUTO. The snapshots are chosen for when the spiral
amplitudes are similar. The agreement between the two codes, as well
as with the 2D simulations, is satisfactory.

%but a longer PLUTO simulation was 
%needed to achieve similar mode amplitudes in ZEUS-MP.

\begin{figure}
  \begin{center}
    \subfigure[ZEUS-MP]{
      \includegraphics[scale=0.33]{figures/polarxy_dens015_zeus}
    }
    \subfigure[PLUTO]{
      \includegraphics[scale=0.33]{figures/pdisk_020}
    }
  \end{center}
  \caption{Three-dimensional simulations using the (a) ZEUS-MP and (b)
    PLUTO. The non-axisymmetric midplane density is shown. Here $\psi
    \equiv \pi/2 - \theta$ is the angular height 
    from the midplane.\label{3d_prelim}}   
\end{figure}

\subsection{Vertical motion}

\subsection{Angular momentum evolution}

@@

@@
